---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sankeyD3plus)
library(tibble)
```

# Smallest Example

The absolute minimum needed to create a Sankey diagram is a **links** data frame with one row per link and columns that define the **source** and **target** as 0-based integer values. By default, all links will be of width 1.

```{r}
links <- data.frame(
  source = c(0, 0, 0, 1, 2),
  target = c(1, 2, 3, 4, 4)
)

sankeyNetwork(
  Links = links,
  Source = "source",
  Target = "target"
)
```

# Labels and Values

One may also provide 

 - a **nodes** data frame to e.g. have proper labels for each node.
 - **values** in the links data frame that define the width of each link.
 - a **unit** within `sankeyNetwork()` which is only visible when hovering over a link or node in the resulting diagram.
 
We will from now on create [tibbles](https://tibble.tidyverse.org/) via [`tribble()`](https://tibble.tidyverse.org/reference/tribble.html) instead of [`data.frames()`](https://rdrr.io/r/base/data.frame.html) for better readability.

```{r}
links <- tribble(
  ~source, ~target, ~value,
  0,       1,       3.7,
  0,       2,       22.9,
  0,       3,       57.7,
  1,       4,       3.7,
  2,       4,       22.9
)

nodes <- tribble(
  ~id, ~label,
  0,   "German residents",
  1,   "Berlin Residents",
  2,   "Other cities (>100k)",
  3,   "Non-Metropolitans",
  4,   "Metropolitans (>100k)"
)

sankeyNetwork(
  Links = links,
  Source = "source",
  Target = "target",
  Value = "value",
  units = "mio. people",
  Nodes = nodes,
  NodeID = "label"
)
```

# Link color & Node color

You can have link-specific and node-specific colors:

```{r}
links <- tribble(
  ~source, ~target, ~value, ~linkcolor,
  0,       1,       3.7,    "#e56b6f",
  0,       2,       22.9,   "#e56b6f",
  0,       3,       57.7,   "#355070",
  1,       4,       3.7,    "#e56b6f",
  2,       4,       22.9,   "#e56b6f"
)

nodes <- tribble(
  ~id, ~label,                  ~nodecolor,
  0,   "German residents",      "#6d597a",
  1,   "Berlin Residents",      "#eaac8b",
  2,   "Other cities (>100k)",  "#b56576",
  3,   "Non-Metropolitans",     "#355070",
  4,   "Metropolitans (>100k)", "#e56b6f"
)

sankeyNetwork(
  Links = links,
  Source = "source",
  Target = "target",
  Value = "value",
  linkColor = "linkcolor",
  Nodes = nodes,
  NodeID = "label",
  NodeColor = "nodecolor",
  units = "mio. people",
)
```

# Node Positioning

You can enforce the vertical order of nodes via `NodePosY` and their horizontal position via `NodePosX`:

```{r}
links <- tribble(
  ~source, ~target, ~value,
  0,       1,       3.7,
  0,       2,       22.9,
  0,       3,       57.7,
  1,       4,       3.7,
  2,       4,       22.9
)

nodes <- tribble(
  ~id, ~label,                  ~yorder, ~xpos,
  0,   "German residents",      1,       0,
  1,   "Berlin Residents",      1,       1,
  2,   "Other cities (>100k)",  3,       1,
  3,   "Non-Metropolitans",     2,       2,
  4,   "Metropolitans (>100k)", 1,       2     
)

sankeyNetwork(
  Links = links,
  Source = "source",
  Target = "target",
  Value = "value",
  units = "mio. people",
  Nodes = nodes,
  NodeID = "label",
  NodePosY = "yorder"
)
sankeyNetwork(
  Links = links,
  Source = "source",
  Target = "target",
  Value = "value",
  units = "mio. people",
  Nodes = nodes,
  NodeID = "label",
  NodePosX = "xpos", # only works in combination with align = "none" or align = "left"
)
```
